The effects of binarity on planet occurrence rates measured by transit surveys

### 2017-11-30 sandbox/note_km.tex
corrected math error in I_1(r_a) model 2

In the presence of dilution, planets with $(\pp\a, \ps\a)$ are associated with 
systems of many different planetary and stellar properties, so $N_{\rm 
det}(\pp\a, \ps\a)$ is 
given by the convolution of the true occurrence, $\Lambda(\pp, \ps)$, and 
number of searchable stars that give $(\pp\a, \ps\a)$ when the true system 
actually has $(\pp, \ps)$, $\mathcal{N}(\pp\a, \ps\a; \pp, \ps)$:
\begin{equation}
	N_{\rm det}(\pp\a, \ps\a)=\sum_i N_{\rm det}^i(\pp\a, \ps\a)
	=\sum_i \int \mathrm{d}\pp\mathrm{d}\ps\,
	%\mathcal{P}_{\rm tra}(\pp\a, \ps\a; \pt)
	\mathcal{N}_{\rm s}^i(\pp\a,\ps\a; \pp, \ps)
	\cdot\Lambda^i(\pp, \ps)\cdot p_{\rm tra}(\pp, \ps),
\end{equation}
where $i$ specifies the type of true host stars (0: single, 1: primary, 2: 
secondary).

@@

Imagine instead that we took $\Lambda_i(r,M) = Z_i r^\delta$.
From Eq.~\ref{eq:apparent_rate_twin_binary}, we get an apparent rate
\begin{equation}
\Lambda_a(r_a) = r_a^\delta \left[
\frac{Z_0}{1+\mu({\rm BF})}
+
2^{\frac{\delta}{2} + 1 } \frac{\mu({\rm BF})}{1+\mu({\rm BF})} \left(
Z_1 + Z_2
\right)
\right].
\end{equation}

Consider the case when the $Z_i$'s are identical. If we compare 
$\Lambda_a(r_a)$ with $\Lambda(r)$, we must impose that $r_a$ and $r$ are 
indistinguishable (i.e.: the observer cannot tell them apart).
Note that the true rate is
\begin{equation}
\Lambda(r) = r^\delta \left[
\frac{Z_0}{1+2\mu} + \frac{\mu}{1+2\mu} (Z_1+Z_2)
\right]
\end{equation}
We then get a correction factor
\begin{equation}
X_\Lambda \equiv \left. \frac{\Lambda_a(r_a)}{\Lambda(r)} 
\right|_{r_a\rightarrow r}
=
\frac{2^{\frac{\delta}{2}+2}\mu + 1}{\mu + 1}.
\end{equation}

@@

@@

\subsubsection{Same Planets (Model \# 2)}
For our ``Model \#2'', we again want $\Lambda_i(r) = Z_i \cdot \delta(r-r_0)$.
To achieve this, write
\begin{equation}
f(q) = \frac{q^\beta (1+q^\alpha)^{3/2}}{\mathcal{N}_q},
\end{equation}
for $\mathcal{N}_q$ the normalization.
For $i=1$, call the integral on the left the $[\ldots]$ ``$I_1$''.
Then
\begin{align}
I_1(r_a) &\equiv \int \frac{{\rm d}q}{\mathcal{A}(q)} f(q) 
    \Lambda_1\left( \frac{r\a}{\mathcal{A}(q)} \right)
    \\
&= \frac{1}{\mathcal{N}_q} \int {\rm d}q\, q^\beta (1+q^\alpha)^2
    \Lambda_1(r_a (1+q^\alpha)^{1/2})
    \\
&= \frac{Z_1}{\mathcal{N}_q} \int {\rm d}q\, q^\beta (1+q^\alpha)^2
    \delta(r_a (1+q^\alpha)^{1/2} - r_0) \\
&= \frac{Z_1}{\mathcal{N}_q}
    \left( \left(\frac{r_0}{r_a}\right)^2 - 1  \right)^{\beta/\alpha}
    \left( \frac{r_0}{r_a}\right)^4.
    \label{eq:model_2_i_1}
\end{align}
The penultimate line used the fact that $r_a(1+q^\alpha)^{1/2}=r_0$ has only 
one root:
\begin{equation}
q(r_a) = \left(\left(\frac{r_0}{r_a}\right)^2 - 1\right)^{1/\alpha}.
\end{equation}
Eq.~\ref{eq:model_2_i_1} should be used as a sanity check on any 
Monte Carlo numerics.

Attempting the same thing for $i=2$,
\begin{align}
I_2(r_a) &\equiv \int \frac{q {\rm d}q}{\mathcal{B}(q)} f(q)
\Lambda_2\left(\frac{r\a}{\mathcal{B}(q)}\right) q^{2/3}
\\
&= \frac{Z_2}{\mathcal{N}_q} \int {\rm d}q\,
    q^{\beta+\frac{8}{3}} 
    (1+q^\alpha)^{3/2}
    (1+q^{-\alpha})^{1/2}
    \delta(r_a q (1+q^{-\alpha})^{1/2} - r_0).
\end{align}
Unfortunately, analytic roots of $r_a q (1+q^{-\alpha})^{1/2} = r_0$ do not 
exist.
Schematically, since for any function $g(x)$,
\begin{equation}
\delta(g(x)) = \sum_{x_i \in \{{\rm roots}\}} \frac{\delta(x-x_i)}{|g'(x_i)|},
\end{equation}
if the roots are found numerically then
\begin{equation}
I_2(r_a) = \frac{Z_2}{\mathcal{N}_q}
    \sum_{q_i \in \{{\rm roots}\}}
q_i^{\beta+\frac{8}{3}} 
(1+q_i^\alpha)^{3/2}
(1+q_i^{-\alpha})^{1/2}.
\end{equation}
But this doesn't really give much insight.

>>>>

In the presence of dilution, planets with $(\pp\a, \ps\a)$ are associated with 
systems of many different planetary and stellar properties, so $N_{\rm 
det}(\pp\a, \ps\a)$ is 
given by the convolution of the true occurrence, $\Lambda(\pp, \ps)$, and 
$\mathcal{N}(\pp\a, \ps\a; \pp, \ps)$, the number of searchable stars that 
give $(\pp\a, \ps\a)$ when the true system actually has $(\pp, \ps)$:
\begin{equation}
	N_{\rm det}(\pp\a, \ps\a)=\sum_i N_{\rm det}^i(\pp\a, \ps\a)
	=\sum_i \int \mathrm{d}\pp\mathrm{d}\ps\,
	%\mathcal{P}_{\rm tra}(\pp\a, \ps\a; \pt)
	\mathcal{N}_{\rm s}^i(\pp\a,\ps\a; \pp, \ps)
	\cdot\Lambda^i(\pp, \ps)\cdot p_{\rm tra}(\pp, \ps),
\end{equation}
where $i$ specifies the type of true host stars (0: single, 1: primary, 2: 
secondary).

@@

Imagine instead that we took $\Lambda_i(r,M) = Z_i f(r) = Z_i 
r^\delta/\mathcal{N}_r$, for $f(r)$ the radius shape function, 
$\mathcal{N}_r$ the shape function's normalization, and $Z_i$ the number of 
planets per star.
From Eq.~\ref{eq:apparent_rate_twin_binary}, we get an apparent rate
\begin{equation}
\Lambda_a(r_a) = \frac{r_a^\delta}{\mathcal{N}_r} \left[
\frac{Z_0}{1+\mu({\rm BF})}
+
2^{\frac{\delta+1}{2} } \frac{\mu({\rm BF})}{1+\mu({\rm BF})} \left(Z_1 + Z_2
\right)
\right].
\end{equation}

Consider the case when the $Z_i$'s are identical. If we compare 
$\Lambda_a(r_a)$ with $\Lambda(r)$, we must impose that $r_a$ and $r$ are 
indistinguishable (i.e.: the observer cannot tell them apart).
Note that the true rate is
\begin{equation}
\Lambda(r) = \frac{r^\delta}{\mathcal{N}_r} \left[
\frac{Z_0}{1+2\mu} + \frac{\mu}{1+2\mu} (Z_1+Z_2)
\right]
\end{equation}
We then get a correction factor
\begin{equation}
X_\Lambda \equiv \left. \frac{\Lambda_a(r_a)}{\Lambda(r)} 
\right|_{r_a\rightarrow r}
=
\frac{2^{\frac{\delta}{2}+2}\mu + 1}{\mu + 1}.
\end{equation}

@@

As a side-note, this particular model (twin binaries, power law planet 
distribution) is a good unit test for the numerics. As of git commit 
\texttt{9b8f55c}, the two agree.

@@

\subsubsection{Same Planets (Model \# 2)}
For our ``Model \#2'', we again want $\Lambda_i(r) = Z_i \cdot \delta(r-r_0)$.
To achieve this, write
\begin{equation}
f(q) = \frac{q^\beta (1+q^\alpha)^{3/2}}{\mathcal{N}_q},
\end{equation}
for $\mathcal{N}_q$ the normalization.
For $i=1$, call the integral on the left the $[\ldots]$ ``$I_1$''.
Then
\begin{align}
I_1(r_a) &\equiv \int \frac{{\rm d}q}{\mathcal{A}(q)} f(q) 
    \Lambda_1\left( \frac{r\a}{\mathcal{A}(q)} \right)
    \\
&= \frac{1}{\mathcal{N}_q} \int {\rm d}q\, q^\beta (1+q^\alpha)^2
    \Lambda_1(r_a (1+q^\alpha)^{1/2})
    \\
&= \frac{Z_1}{\mathcal{N}_q} \int {\rm d}q\, q^\beta (1+q^\alpha)^2
    \delta(r_a (1+q^\alpha)^{1/2} - r_0) \\
&= \frac{Z_1}{\mathcal{N}_q} \frac{2}{\alpha} \frac{r_0}{r_a^2} 
   \left(\left(\frac{r_0}{r_a}\right)^2 -1
       \right)^{\frac{1-\alpha}{\alpha}}
   \int {\rm d}q\, q^\beta (1+q^\alpha)^2 
       \delta\left( q - \left( \left(\frac{r_0}{r_a}\right)^2 -1
       \right)^{\frac{1}{\alpha}} \right) \\
&= \frac{Z_1}{\mathcal{N}_q} \frac{2}{\alpha} \frac{r_0}{r_a^2}
    \left( \left(\frac{r_0}{r_a}\right)^2 - 1  
    \right)^{\frac{\beta-\alpha+1}{\alpha}}
    \left( \frac{r_0}{r_a}\right)^4.
    \label{eq:model_2_i_1}
\end{align}
The penultimate line used the fact that $r_a(1+q^\alpha)^{1/2}=r_0$ has only 
one root:
\begin{equation}
q(r_a) = \left(\left(\frac{r_0}{r_a}\right)^2 - 1\right)^{1/\alpha},
\end{equation}
and for any function $f(q)$ with a single root $q_i$,
\begin{equation}
\delta(f(q)) = \frac{\delta(q-q_i)}{|f'(q_i)|}.
\end{equation}
Eq.~\ref{eq:model_2_i_1} is a good unit test for Monte Carlo numerics.

Attempting the same thing for $i=2$,
\begin{align}
I_2(r_a) &\equiv \int \frac{q {\rm d}q}{\mathcal{B}(q)} f(q)
\Lambda_2\left(\frac{r\a}{\mathcal{B}(q)}\right) q^{2/3}
\\
&= \frac{Z_2}{\mathcal{N}_q} \int {\rm d}q\,
    q^{\beta+\frac{8}{3}} 
    (1+q^\alpha)^{3/2}
    (1+q^{-\alpha})^{1/2}
    \delta(r_a q (1+q^{-\alpha})^{1/2} - r_0).
\end{align}
Unfortunately, analytic roots of $r_a q (1+q^{-\alpha})^{1/2} = r_0$ do not 
exist.
If the roots are found numerically then
\begin{equation}
I_2(r_a) = \frac{Z_2}{\mathcal{N}_q}
    \sum_{q_i \in \{{\rm roots}\}}
q_i^{\beta+\frac{8}{3}} 
(1+q_i^\alpha)^{3/2}
(1+q_i^{-\alpha})^{1/2},
\end{equation}
but this doesn't really give much insight.
